### 2022-06-22 manuscript/tables/notebook_overview.tex
minor fix

\begin{tabular}{p{0.25\textwidth}p{0.55\textwidth}p{0.1\textwidth}}
    \hline
    \textbf{Topic} & \textbf{Description} & \textbf{Hyperlink} \\
    \hline
    \hline
    What is a kinase? & Introduction to kinases and challenges in drug discovery. & \href{https://projects.volkamerlab.org/teachopencadd/talktorials/T023\_what\_is\_a\_kinase.html}{T023} \\
    Pocket sequence & Pairwise similarities/identities between $85$ residue long KLIFS pocket sequences. & \href{https://projects.volkamerlab.org/teachopencadd/talktorials/T024\_kinase\_similarity\_sequence.html}{T024} \\
    Pocket structure & Pairwise similarities between $1,032-$bit long KiSSim fingerprints, which encode spatial and physico-chemical pocket properties. & \href{https://projects.volkamerlab.org/teachopencadd/talktorials/T025\_kinase\_similarity\_kissim.html}{T025} \\
    Pocket-ligand interactions & Pairwise similarities between $595-$bit long KLIFS kinase-ligand interaction fingerprints (IFP). & \href{https://projects.volkamerlab.org/teachopencadd/talktorials/T026\_kinase\_similarity\_ifp.html}{T026} \\
    Ligand profile & Pairwise similarity based on the ratio of compounds tested active against kinase pairs.& \href{https://projects.volkamerlab.org/teachopencadd/talktorials/T027\_kinase\_similarity\_ligand\_profile.html}{T027} \\
    Kinase similarity & Comparison between predicted off-targets based on calculated kinase similarities using aforementioned methods. & \href{https://projects.volkamerlab.org/teachopencadd/talktorials/T028\_kinase\_similarity\_compare\_perspectives.html}{T028} \\
    \hline
\end{tabular}